SIR Model With Testing Flows and Isolation

### 2020-12-01 codes/ali_test_sir.Rmd
panel plot updated

- $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively. We assume that $\eta_t<\eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
- $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.  It has units of $1/time$.
- $\gamma$: recovery rate ($1/time$).
- $\rho$: per capita testing intensity across the whole population ($1/time$).
- $W$: weighted number of people available for tests, defined as $W = W_S S_u + W_I I_u + W_R R_u$.
- $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\rho N_0}{W}$.
- $F_Z$: Weighted testing rate defined as $F_Z=\sigma W_Z$. That is, $F_S = \sigma W_S$, $F_I=\sigma W_I$ and $F_R = \sigma W_R$.

@@

```{r, model formulation}
sir.model <- function(time,state,params){
    unpack(as.list(c(state,params)))
    ## Force of Infection  
    Lambda <- beta * (I_u + eta_w*(I_n+I_p) + eta_t*I_t)/N0
    ## scaling the weights
    W <- W_S*S_u+W_I*I_u+W_R*R_u
    sigma <- rho*N0/W
    F_S <- sigma*W_S
    F_I <- sigma*W_I
    F_R <- sigma*W_R

@@

```{r model inputs, message=FALSE}
params <- c(N0=1000000, beta=.8,gamma=1/3, omega=0.25, rho=0.01, #0.8, 1/3
             W_S=1, W_I=1, W_R=1, #weights
             p_S=0, p_I=1, p_R=0.5, #prob of waiting for being tested positive
            eta_w=0.02, eta_t=0.01, #isolation parameter 
            s=2 #eta_w/eta_t
            )
class(params) <- "params_pansim"
state_init <- c(S_u=params[["N0"]], S_n=0,
                I_u=1,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)
d <- seq(0,200,by=0.1)
# Conditional stop for desolver
# rootfun <- function (time, state,params) { 
#   unpack(as.list(c(state,params)))
#   return(I_u - 1) }
```

@@

$$R_0= (A \times S_u^* + B \times S_n^*) \times C,$$ where
$A=\gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t p_I) F_I$,
$B=\big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega p_I F_I }{\omega+\gamma}$ and 
$C=\frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}$.

\begin{equation}\label{FV}
F= \beta/N_0 \left[ \begin {array}{cccc} 
S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_t\,S_u\\
S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_t\,S_n\\ 
0&0&0&0\\
0&0&0&0
 \end {array} \right],
 V=
 \left[ \begin {array}{cccc}  
F_I+\gamma&-\omega&0&0\\
-(1-p_I)F_I&\omega+\gamma&0&0\\
-p_I F_I&0&\omega+\gamma&0\\
0&0&-\omega&\gamma
\end {array} \right],
\end{equation}

@@

R0<-function(params){
    unpack(as.list(params))
    Sn <- Sn_dfe(params)
    Su <- N0-Sn
    # print(c(Su_dfe,Sn_dfe,Su_dfe+Sn_dfe))
    Fi <- F_I(params) #at DFE
    A <-gamma*(omega+gamma+eta_w*Fi) + omega*eta_t*p_I*Fi
    B <-(omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_t*omega)*(omega*p_I*Fi)/(omega+gamma)
    C <- beta/(N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*p_I)))
    return((A*Su+B*Sn)*C)
}

# R0 expression with eta_w/eta_t factored
R02<-function(params){
    unpack(as.list(params))
    Sn<- Sn_dfe(params)
    Su<- N0-Sn
    F_I <-F_I(params)  #at DFE

    A2 <- gamma*(p_I*F_I*omega*Sn+(omega+gamma)*(gamma*Sn+F_I*N0))
    B2 <- p_I*F_I*omega*(gamma*Su+omega*N0)
    D <- gamma*(omega+gamma)*(gamma*Su+omega*N0)
    C2 <- beta/((N0*gamma*(gamma*(omega+gamma)+F_I*(gamma+omega*p_I)))*(omega+gamma))
    # s <- eta_w/eta_t
    return((eta_t*(s*A2+B2)+D)*C2)
}

@@

Can we recover $R_0=\beta/\gamma$ of the simple SIR? -yes
```{r checkings}
# checking if simple R_0 for SIR can be recovered?
# Case #1: rho=0
all.equal(R0(params=update(params,rho=0)), params[["beta"]]/params[["gamma"]])
# Case #2: 
all.equal(R0(params=update(params,eta_t=1,eta_w=1)), params[["beta"]]/params[["gamma"]])
```

### Contour plot of $R_0=1$ and interpretations approache 1
```{r contour_R0_1, message=FALSE,warning=FALSE }
n_out <- 5 ## facets (rows/cols)
n_in <- 41 ## grid N within facets
 # specify the ranges
 beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case
 eta_w <- seq(0,1, length.out=n_out)
 eta_t <- seq(0,1, length.out=n_out) #0.001 to 0.5 
 rho     <- seq(0,0.01, length.out=n_in)
 omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.011
 W_S <- 0.3 #choose 0.3 for TTI testing plan and 1 for the random testing.

 df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                    W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
                    eta_w=eta_w,eta_t=eta_t,s=params[["s"]])

@@

df2 <- df2 %>% dplyr::mutate(R0_sub = ifelse(eta_w<eta_t, NA, R0))

which(df2[]<0)
```

@@

```{r show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
p1 <- ggplot(df2,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
  xlab(TeX('$\\omega$, Returning test result rate (1/day)')) +
  ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
    theme(panel.spacing=grid::unit(0,"lines"))
p1_temp <- (p1
    + geom_contour_filled(breaks=pretty(unique(df2$R0_sub),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=1,alpha=0.5,colour="black")
    + facet_grid(eta_w~eta_t, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="R[0]"))
    + ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S)))
)
p1_temp
# ggsave(p1_temp,filename = "R0contour_random.png" , path="../pix/")
ggsave(p1_temp,filename = "R0contour_TTI.png" , path="../pix/")

@@

### Contour plot of $R_0=1$ and interpretations approache 2, parametrization $s=\eta_w/\eta_t$

```{r contour_R0_2, message=FALSE,warning=FALSE }
 # specify the ranges
beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case 
rho     <- seq(0,0.01, length.out=n_in)
omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, 
eta_t <- seq(0.001,0.1, length.out=n_out)
s <- seq(1,10,length.out=n_out)

 df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                    W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s)

@@

```{r show_contours_2, width=24,height=24,message=FALSE,warning=FALSE}
p11 <- ggplot(df22,aes(x=omega,y=rho,z=R0))+ theme_bw() +
    theme(panel.spacing=grid::unit(0,"lines"))
p2_temp <- (p11
    + geom_contour_filled(breaks=pretty(unique(df22$R0),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=1,alpha=0.5,colour="black")
    + facet_grid(s~eta_t, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="R[0]"))
    + ggtitle("W[S]=1, W[I]=W[R]=1")
)
p2_temp
# ggsave(p2_temp,filename = "R0contour_random_s.png" , path="../pix/")
```

@@

1. $\partial{R_0}/\partial{\eta_t} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$.
That is, the lower isolation the higher $R_0$. Note that perfect isolation occurs at $\eta_w = 0$. 
2. $\partial{R_0}/\partial{\rho} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $\rho=0$ which is
$R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2)$.
3. Based on the above Taylor expansion around $\rho=0$,
$$\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
where $a=(\eta_w-1)W_I-(\eta_t-1)P_I W_I = ((s-p_I)\eta_t + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
Given that $\eta_t\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$.
Let $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$. Thus, $\partial{R_0}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{R_0}/\partial{\omega}<0$ for $\omega>\omega_2$.

In general, $a \geq 0$ is not true. $$a \geq 0 \iff (s-p_I) \eta_t \geq (1-p_I),$$ where $s=\frac{\eta_w}{\eta_t} \geq 1$. As an example, in the case of a very accurate testing regime,  i.e., $P_I=1$, $a \geq 0$ is achieved.

**Perfect isolation**; When $\eta_w=0$ and consequently $\eta_t=0$. This leads to $a \leq 0$, $b<0$, $c<0$, thus $\partial{R_0}/\partial{\omega} \geq 0$. Intuitively, returning the test results back to the awaiting individuals while they are in a perfect isolation results in an increase of $R_0$.

@@

(i) in the case of perfect isolation of the reported individuals, $\eta_t=0$, the followings are inferred (see the first few columns of the panel plot); 
 - variations of the $\eta_w$ has a negligible effect on the critical testing.
 - for low $\omega$, $R_0$ increases as $\omega$ increases but saturates. This means that when the reporting process is very very slow, increasing the speed of the reporting causes higher $R_0$. Maybe because more people will move to $I_u$ [?, not very sure here.]
 (ii) in case of the non-perfect isolation (see the last column), when $\eta_w$ increases (more awaiting people follow the isolation rules) faster reporting reduces $R_0$.  
Note that it is sensible to expect that $\partial{R_0}/\partial{\omega}<0$. Here we showed that this can be less true.

@@

```{r  chnages of R0 wrt omega,message=FALSE,warning=FALSE }
R0_omega <- function(params){
  unpack(as.list(params))
         # a <- (eta_w-1)*W_I-(eta_t-1)*p_I*W_I
         a <- ((s-p_I)*eta_t + (p_I-1))*W_I  #involving s
         b <- 2*(eta_w-1)*gamma*W_S
         c <- b*gamma/2

@@

# eta_w <- seq(0,1, length.out=n_out)
eta_t <- seq(0.001,1, length.out=n_out)
s <- seq(1,2,length.out=n_out)

@@

 df3_1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=gamma,omega=omega,rho=rho,
                    W_S=W_S,W_I=W_I,W_R=W_R,
                    p_S=params[["p_S"]],p_I=p_I,p_R=params[["p_R"]],
                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s)

@@

```{r show_contours_3, width=24,height=24,message=FALSE,warning=FALSE}
if(F){
p3 <- ggplot(df3_2,aes(x=omega,y=gamma,z=R0_omega))+ theme_bw() +
    theme(panel.spacing=grid::unit(0,"lines"))
(p3
    + geom_contour_filled(breaks=pretty(unique(df3_2$R0_omega),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=0,alpha=0.5,colour="black")
    + facet_grid(s~eta_t, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="dR[0]"))
)
}
```

>>>>

- $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_c I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_c$ are the isolation parameters for awaiting and reported individuals, respectively. We assume that $\eta_c<\eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
- $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.  It has units of $1/time$.
- $\gamma$: recovery rate ($1/time$).
- $\rho$: per capita testing intensity across the whole population ($1/time$).
- $W$: weighted number of people available for tests, defined as $W = W_S S_u + W_I I_u + W_R R_u$.
- $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\rho N_0}{W}$.
- $F_Z$: Weighted testing rate defined as $F_Z=\sigma W_Z$. That is, $F_S = \sigma W_S$, $F_I=\sigma W_I$ and $F_R = \sigma W_R$.

@@

```{r, model formulation}
sir.model <- function(time,state,params){
    unpack(as.list(c(state,params)))
    ## Force of Infection  
    Lambda <- beta * (I_u + eta_w*(I_n+I_p) + eta_c*I_t)/N0
    ## scaling the weights
    W <- W_S*S_u+W_I*I_u+W_R*R_u
    sigma <- rho*N0/W
    F_S <- sigma*W_S
    F_I <- sigma*W_I
    F_R <- sigma*W_R

@@

```{r model inputs, message=FALSE}
params <- c(N0=1000000, beta=.8,gamma=1/3, omega=0.25, rho=0.01, #0.8, 1/3
             W_S=1, W_I=1, W_R=1, #weights
             p_S=0, p_I=1, p_R=0.5, #prob of waiting for being tested positive
            eta_w=0.02, eta_c=0.01, #isolation parameter 
            s=2 #eta_w/eta_c
            )
class(params) <- "params_pansim"
state_init <- c(S_u=params[["N0"]], S_n=0,
                I_u=1,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)
d <- seq(0,200,by=0.1)
# Conditional stop for desolver
# rootfun <- function (time, state,params) { 
#   unpack(as.list(c(state,params)))
#   return(I_u - 1) }
```

@@

$$R_0= (A \times S_u^* + B \times S_n^*) \times C,$$ where
$A=\gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_c p_I) F_I$,
$B=\big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_c\omega) \omega p_I F_I }{\omega+\gamma}$ and 
$C=\frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}$.

\begin{equation}\label{FV}
F= \beta/N_0 \left[ \begin {array}{cccc} 
S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_c\,S_u\\
S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_c\,S_n\\ 
0&0&0&0\\
0&0&0&0
 \end {array} \right],
 V=
 \left[ \begin {array}{cccc}  
F_I+\gamma&-\omega&0&0\\
-(1-p_I)F_I&\omega+\gamma&0&0\\
-p_I F_I&0&\omega+\gamma&0\\
0&0&-\omega&\gamma
\end {array} \right],
\end{equation}

@@

R0<-function(params){
    unpack(as.list(params))
    Sn <- Sn_dfe(params)
    Su <- N0-Sn
    # print(c(Su_dfe,Sn_dfe,Su_dfe+Sn_dfe))
    Fi <- F_I(params) #at DFE
    A <-gamma*(omega+gamma+eta_w*Fi) + omega*eta_c*p_I*Fi
    B <-(omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_c*omega)*(omega*p_I*Fi)/(omega+gamma)
    C <- beta/(N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*p_I)))
    return((A*Su+B*Sn)*C)
}

# R0 expression with eta_w/eta_c factored
R02<-function(params){
    unpack(as.list(params))
    Sn<- Sn_dfe(params)
    Su<- N0-Sn
    F_I <-F_I(params)  #at DFE

    A2 <- gamma*(p_I*F_I*omega*Sn+(omega+gamma)*(gamma*Sn+F_I*N0))
    B2 <- p_I*F_I*omega*(gamma*Su+omega*N0)
    D <- gamma*(omega+gamma)*(gamma*Su+omega*N0)
    C2 <- beta/((N0*gamma*(gamma*(omega+gamma)+F_I*(gamma+omega*p_I)))*(omega+gamma))
    # s <- eta_w/eta_c
    return((eta_c*(s*A2+B2)+D)*C2)
}

@@

Can we recover $R_0=\beta/\gamma$ of the simple SIR? -yes
```{r checkings}
# checking if simple R_0 for SIR can be recovered?
# Case #1: rho=0
all.equal(R0(params=update(params,rho=0)), params[["beta"]]/params[["gamma"]])
# Case #2: 
all.equal(R0(params=update(params,eta_c=1,eta_w=1)), params[["beta"]]/params[["gamma"]])
```

### Contour plot of $R_0=1$ and interpretations approache 1
```{r contour_R0_1, message=FALSE,warning=FALSE }
n_out <- 5 ## facets (rows/cols)
n_in <- 41 ## grid N within facets
 # specify the ranges
 beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case
 eta_w <- seq(0,1, length.out=n_out)
 eta_c <- seq(0,1, length.out=n_out) #0.001 to 0.5 
 rho     <- seq(0,0.01, length.out=n_in)
 omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, was 0.011
 W_S <- 1 #choose 0.3 for TTI testing plan and 1 for the random testing.

 df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                    W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
                    eta_w=eta_w,eta_c=eta_c,s=params[["s"]])

@@

df2 <- df2 %>% dplyr::mutate(R0_sub = ifelse(eta_w<eta_c, NA, R0))

# df2_fake <- df2 %>% dplyr::select(eta_c,eta_w)
# df2_fake <- unique(df2_fake) %>% dplyr::mutate(flag=ifelse(eta_c<=eta_w,1,0))

@@

```{r show_contours_1, width=24,height=24,message=FALSE,warning=FALSE}
p1 <- ggplot(df2,aes(x=omega,y=rho,z=R0_sub))+ theme_bw() +
  xlab(TeX('$\\omega$, Returning test result rate (1/day)')) +
  ylab(TeX('$\\rho$, testing intensity (1/day per capita)')) +
    # theme(panel.spacing=grid::unit(0,"lines")) # initial theme, comment it for TTI testing plot
    theme(panel.spacing=grid::unit(0,"lines"),legend.position = "none") #AG, uncomment for Random Case
# p1_temp <- (p1
#     + geom_contour_filled(breaks=pretty(unique(df2$R0_sub),5),
#                           alpha=0.5) ## breaks=0.2)
#     + geom_contour(breaks=1,alpha=0.5,colour="black")
#     + facet_grid(eta_w~eta_c, labeller=label_special)
#     + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
#     + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
#     + scale_fill_viridis_d(name=parse(text="R[0]"))
#     + ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S)))
#     + geom_rect(data=df2,
#                           fill=ifelse(df2$eta_c>df2$eta_w,"white","NA" ),
#                           color= NA,
#                           ymin=-1,
#                           ymax=10,
#                           xmin=-1,
#                           xmax=2)
#     )

p1_temp <- (p1
    + geom_contour_filled(breaks=seq(0.80,1.05,by=0.05)
                          )
    + geom_contour(breaks=1,alpha=0.5,colour="black")
    + facet_grid(eta_w~eta_c, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="R[0]"))
    + ggtitle(TeX(sprintf("W_S=%.1f, W_I=W_R=1",W_S)))
    + geom_rect(data=df2,
                          fill=ifelse(df2$eta_c>df2$eta_w,"white","NA" ),
                          color= NA,
                          ymin=-1,
                          ymax=10,
                          xmin=-1,
                          xmax=2)
    )
p1_temp
# ## hack (By Dr. Ben Bolker) white rectangle to cover grid
#   ggplot() + 
#   scale_x_continuous(name="x") + 
#   scale_y_continuous(name="y") +
#   facet_grid(eta_w~eta_c, labeller=label_special) +
#     geom_rect(data=df2_fake,
#                           fill=ifelse(df2_fake$eta_c>df2_fake$eta_w,"white","black"),
#                           color= NA,
#                           ymin=-1,
#                           ymax=10,
#                           xmin=-1,
#                           xmax=2)

ggsave(p1_temp,filename = "R0contour_random.png" , path="../pix/")
# ggsave(p1_temp,filename = "R0contour_TTI.png" , path="../pix/")

@@

### Contour plot of $R_0=1$ and interpretations approache 2, parametrization $s=\eta_w/\eta_c$

```{r contour_R0_2, message=FALSE,warning=FALSE }
 # specify the ranges
beta  <- 0.339  ## set beta high enough to allow R0>1 in worst case 
rho     <- seq(0,0.01, length.out=n_in)
omega <- seq(0.1,2 , length.out=n_in) #note omega must be > rho, 
eta_c <- seq(0.001,0.1, length.out=n_out)
s <- seq(1,10,length.out=n_out)

 df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                    W_S=W_S,W_I=params[["W_I"]],W_R=params[["W_R"]],
                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
                    eta_w=params[["eta_w"]],eta_c=eta_c,s=s)

@@

```{r show_contours_2, width=24,height=24,message=FALSE,warning=FALSE}
p11 <- ggplot(df22,aes(x=omega,y=rho,z=R0))+ theme_bw() +
    theme(panel.spacing=grid::unit(0,"lines"))
p2_temp <- (p11
    + geom_contour_filled(breaks=pretty(unique(df22$R0),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=1,alpha=0.5,colour="black")
    + facet_grid(s~eta_c, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="R[0]"))
    + ggtitle("W[S]=1, W[I]=W[R]=1")
)
p2_temp
# ggsave(p2_temp,filename = "R0contour_random_s.png" , path="../pix/")
```

@@

1. $\partial{R_0}/\partial{\eta_c} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$.
That is, the lower isolation the higher $R_0$. Note that perfect isolation occurs at $\eta_w = 0$. 
2. $\partial{R_0}/\partial{\rho} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $\rho=0$ which is
$R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_c -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2)$.
3. Based on the above Taylor expansion around $\rho=0$,
$$\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
where $a=(\eta_w-1)W_I-(\eta_c-1)P_I W_I = ((s-p_I)\eta_c + (p_I-1)) W_I$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
Given that $\eta_c\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$.
Let $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$. Thus, $\partial{R_0}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{R_0}/\partial{\omega}<0$ for $\omega>\omega_2$.

In general, $a \geq 0$ is not true. $$a \geq 0 \iff (s-p_I) \eta_c \geq (1-p_I),$$ where $s=\frac{\eta_w}{\eta_c} \geq 1$. As an example, in the case of a very accurate testing regime,  i.e., $P_I=1$, $a \geq 0$ is achieved.

**Perfect isolation**; When $\eta_w=0$ and consequently $\eta_c=0$. This leads to $a \leq 0$, $b<0$, $c<0$, thus $\partial{R_0}/\partial{\omega} \geq 0$. Intuitively, returning the test results back to the awaiting individuals while they are in a perfect isolation results in an increase of $R_0$.

@@

(i) in the case of perfect isolation of the reported individuals, $\eta_c=0$, the followings are inferred (see the first few columns of the panel plot); 
 - variations of the $\eta_w$ has a negligible effect on the critical testing.
 - for low $\omega$, $R_0$ increases as $\omega$ increases but saturates. This means that when the reporting process is very very slow, increasing the speed of the reporting causes higher $R_0$. Maybe because more people will move to $I_u$ [?, not very sure here.]
 (ii) in case of the non-perfect isolation (see the last column), when $\eta_w$ increases (more awaiting people follow the isolation rules) faster reporting reduces $R_0$.  
Note that it is sensible to expect that $\partial{R_0}/\partial{\omega}<0$. Here we showed that this can be less true.

@@

```{r  chnages of R0 wrt omega,message=FALSE,warning=FALSE }
R0_omega <- function(params){
  unpack(as.list(params))
         # a <- (eta_w-1)*W_I-(eta_c-1)*p_I*W_I
         a <- ((s-p_I)*eta_c + (p_I-1))*W_I  #involving s
         b <- 2*(eta_w-1)*gamma*W_S
         c <- b*gamma/2

@@

# eta_w <- seq(0,1, length.out=n_out)
eta_c <- seq(0.001,1, length.out=n_out)
s <- seq(1,2,length.out=n_out)

@@

 df3_1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=gamma,omega=omega,rho=rho,
                    W_S=W_S,W_I=W_I,W_R=W_R,
                    p_S=params[["p_S"]],p_I=p_I,p_R=params[["p_R"]],
                    eta_w=params[["eta_w"]],eta_c=eta_c,s=s)

@@

```{r show_contours_3, width=24,height=24,message=FALSE,warning=FALSE}
if(F){
p3 <- ggplot(df3_2,aes(x=omega,y=gamma,z=R0_omega))+ theme_bw() +
    theme(panel.spacing=grid::unit(0,"lines"))
(p3
    + geom_contour_filled(breaks=pretty(unique(df3_2$R0_omega),5),
                          alpha=0.5) ## breaks=0.2)
    + geom_contour(breaks=0,alpha=0.5,colour="black")
    + facet_grid(s~eta_c, labeller=label_special)
    + scale_x_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_y_continuous(expand=expansion(c(0,0)), n.breaks=3)
    + scale_fill_viridis_d(name=parse(text="dR[0]"))
)
}
```
